---
title: "cce_ground_cleaning"
author: "Weronika Konwent"
date: "2024-09-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir="/Users/wk78/Library/CloudStorage/OneDrive-NorthernArizonaUniversity/R/CCE")
library(dplyr)
```

## Cleaning CCE_ground data
#### Here we will clean CCE_ground data and fix any typos, inconsistencies, and other potential issues.


```{r cce_ground}
#read in the datafile
cce_ground <- read.csv("data/working/raw/cce_ground.csv")

#let's have a little preview
head(cce_ground)

#ok, this dataset is pretty big. 
sort(unique(cce_ground$fire_scar))
cce_ground <- cce_ground %>%
  mutate(fire_scar = case_when(fire_scar == "Shovel creek" ~ "Shovel Creek",
                               fire_scar == "Hess creek" ~ "Hess Creek", TRUE ~ fire_scar))

#fix site name typos
sort(unique(cce_ground$site))
cce_ground <- cce_ground %>%
  mutate(site = case_when(site == "MCCR6" ~ "McCr6",
                          site == "HAYS5" ~ "Hays5", 
                          site == "MCCR2" ~ "McCr2",
                          site == "MCCR1" ~ "McCr1",
                          site == "MCCR5" ~ "McCr5", 
                          site == "HAYS6" ~ "Hays6",
                          site == "MCCR4" ~ "McCr4", 
                          site == "HAYS4" ~ "Hays4", TRUE ~ site))

#let's move the one site that has location_other
cce_ground <- cce_ground %>%
  mutate(location = ifelse(location == "" | is.na(location), location_other, location))

#then let's move ground_cover_other to ground_cover
cce_ground <- cce_ground %>%
  mutate(ground_cover_type = ifelse(ground_cover_type == "" | is.na(ground_cover_type), ground_cover_type_other, ground_cover_type))

#there's one site note that changes the location. let's make that change in the location column
cce_ground <- cce_ground %>%
  mutate(location = ifelse(site_notes == "0m, not 9m. Entered as 9 due to fulcrum set up.", "0 m", location))

#perhaps let's add a project column
cce_ground <- cce_ground %>%
  mutate(project = "CCE")

#let's clean up our columns
cce_ground <- select(cce_ground, date, fire_scar, site, plot, location, site_notes, ground_cover_type, ground_percent_cover, ground_cover_notes)

write.csv(cce_ground, file = "data/working/clean/cce_ground.csv")
```

#### These data are now clean and ready for analysis. I removed photos from this clean version, but the photo IDs still exist in the working/raw folder. Percentages per plot can be more than 100%, but just counting trace. For example, the total per plot can be 100.7%, and this would indicate that there were 7 counts of things that were only trace.