---
title: "cce_seedlings_cleaning"
author: "Weronika Konwent"
date: "2024-09-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir="/Users/wk78/Library/CloudStorage/OneDrive-NorthernArizonaUniversity/R/CCE")
library(dplyr)
```

## Cleaning CCE_seedlings data
#### Here we will clean CCE_seedlings data and fix any typos, inconsistencies, and other potential issues.

```{r cce_seedlings}
#read in the datafile
cce_seedlings <- read.csv("data/working/raw/cce_seedlings.csv")

#let's have a little preview
head(cce_seedlings)

#some clean up of sites and fire_scars
sort(unique(cce_seedlings$fire_scar))
cce_seedlings <- cce_seedlings %>%
  mutate(fire_scar = case_when(fire_scar == "Shovel creek" ~ "Shovel Creek",
                               fire_scar == "Hess creek" ~ "Hess Creek", TRUE ~ fire_scar))

#fix site name typos
sort(unique(cce_seedlings$site))
cce_seedlings <- cce_seedlings %>%
  mutate(site = case_when(site == "MCCR6" ~ "McCr6",
                          site == "HAYS5" ~ "Hays5", 
                          site == "MCCR2" ~ "McCr2",
                          site == "MCCR1" ~ "McCr1",
                          site == "MCCR5" ~ "McCr5", 
                          site == "HAYS6" ~ "Hays6",
                          site == "MCCR4" ~ "McCr4", 
                          site == "HAYS4" ~ "Hays4", TRUE ~ site))

#let's check that there's no weirdness with species
sort(unique(cce_seedlings$species_))

#well first step is renaming that column
cce_seedlings <- cce_seedlings %>%
  rename(species = species_)

#let's move the one site that has location_other
cce_seedlings <- cce_seedlings %>%
  mutate(location = ifelse(location == "" | is.na(location), location_other, location))

#there's one site note that changes the location. let's make that change in the location column
cce_seedlings <- cce_seedlings %>%
  mutate(location = ifelse(site_notes == "0m, not 9m. Entered as 9 due to fulcrum set up.", "0 m", location))

#ok, first problem I see is that sometimes a bd was accidentally put in seedling notes. i think the best way to fix this is to paste those numbers in bd3
cce_seedlings <- cce_seedlings %>%
  mutate(bd3 = ifelse(!is.na(seedling_notes) & seedling_notes != "", seedling_notes, bd3))

#next problem is that in HeCr5C 9 we have a seedling count of 0, but with two bds. will change the count here to 2
cce_seedlings <- cce_seedlings %>%
  mutate(seedling_count = ifelse(seedling_count == 0, 2, seedling_count))

###remove comments if we want each row to be it's own observation
#let's mutate the count column so that each row is it's own observation
#cce_seedlings <- cce_seedlings %>%
  #first we have to change all of the blanks to 1s
#  mutate(seedling_count = ifelse(seedling_count == "" | is.na(seedling_count), 1, as.numeric(seedling_count))) %>%
#  uncount(weights = seedling_count)

#perhaps let's add a project column
cce_seedlings <- cce_seedlings %>%
  mutate(project = "CCE")

#let's clean up our columns; there was no rbd3, so I removed it
cce_seedlings <- select(cce_seedlings, date, project, fire_scar, site, plot, location, site_notes, seedlings_present, species, seedling_count, bd1, bd2, bd3)

write.csv(cce_seedlings, file = "data/working/clean/cce_seedlings.csv")
```

#### These data are now clean and ready for analysis. I removed photos from this clean version, but the photo IDs still exist in the working/raw folder. I did not change the count column so that each row is it's own observation. the code for that has been commented out at line 65